Fixing Vulnerabilities Hinders Maintainability

### 2020-01-27 paper/icpc20.tex
more polising on the research questions'

modifying the software code base. However, these refactorings may have a

@@

performant one. Thus, we want to understand how security refactorings impact

@@

\begin{framed}
\textit{\textbf{RQ1} What is the impact of security refactorings on the
maintainability of open-source software?}
\end{framed}
\vspace{-0.1cm}

Often security flaws require refactoring code to make software more secure.
However, there is no evidence yet of how security refactorings impact the
maintainability of open-source software. Our suspicion is that developers tend
to introduce technical debt in their software when refactoring the code base to
address a security flaw because they tend to choose the easiest path to solve
it. To address it, we compute the maintainability of $607$ commits using the
\emph{Better Code Hub} tool. The same approach is applied to a randomly generated
dataset of regular commits --- baseline --- that we use to understand how
maintainability evolves when security refactorings are performed versus when
they are not.

\begin{framed}
\textit{\textbf{RQ2} Which patterns of security refactorings are more likely to
affect open-source software maintainability?}
\end{framed}
\vspace{-0.1cm}
% here
There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a
cross-site scripting vulnerability since usually the last one can be fixed
without adding new lines of code. A typical fix for the cross-site scripting
vulnerability is presented in Listing~\ref{lst:fix}. The developer added the
function \texttt{htmlentities} to \textit{escape} the data given by the variable
\texttt{\$\_['file']}. Knowing which patterns are more likely to increase
maintainability issues is one step forward to bring awareness to security
engineers of what patterns need more attention. The taxonomy of security
patterns used to answer this question is the same as the one provided by the
authors of the dataset~\cite{Reis:2017:IJSSE}. Maintainability is measured
separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
  <li class='error'>
   <?php echo $l->t('Cloud not found');?>
   <p class='hint'>
    <?php
-   if(isset($_['file'])) echo $_['file']
    ?>
   </p>
   <p class='hint'>
    <?php
+   if(isset($_['file'])) echo htmlentities($_['file'])
    ?>
   </p>
  </li>
\end{lstlisting}

>>>>

modifying the software code base. However, these patches may have a

@@

performant one. Thus, we want to understand how security patches impact

@@

\begin{framed}
\textit{\textbf{RQ1} What is the impact of security patches on the
maintainability of open-source software?}
\end{framed}
\vspace{-0.1cm}

Often security flaws require patching code to make software more secure.
However, \textbf{there is no evidence yet of how security patches impact the
maintainability of open-source software}. Our suspicion is that developers tend
to introduce technical debt in their software when patching the code base to
address a security flaw because they tend to choose the easiest path to solve
it. To address it, we compute the maintainability of $1037$ commits using the
\emph{Better Code Hub} tool. The same approach is applied to a randomly generated
dataset of regular commits --- baseline --- that we use to understand how
maintainability evolves when security patches are performed versus when
they are not.

\begin{framed}
\textit{\textbf{RQ2} Which patterns of security patches are more likely to
affect open-source software maintainability?}
\end{framed}
\vspace{-0.1cm}
% here
There are security flaws that are more difficult to patch than others. For
instance, implementing secure authentication is not as easy as fixing a
cross-site scripting vulnerability since usually the last one can be fixed
without adding new lines of code/complexity to the code. A typical fix for 
the cross-site scripting vulnerability is presented in Listing~\ref{lst:fix}. The developer added the function \texttt{htmlentities} to \textit{escape} the data given by the variable
\texttt{\$\_['file']}. \textbf{Knowing which patterns are more likely to increase
maintainability issues is one step forward to bring awareness to security
engineers of what patterns need more attention}. The taxonomy of security
patterns used to answer this question is the one provided by the Common Weakness Enumeration
(CWE). Maintainability is measured
separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
   <p class='hint'>
    <?php
-   if(isset($_['file'])) echo $_['file']
    ?>
   </p>
   <p class='hint'>
    <?php
+   if(isset($_['file'])) echo htmlentities($_['file'])
    ?>
   </p>
\end{lstlisting}
